---
title: "final_R_reference"
author: "Ragib Ahsan"
date: "12/16/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages

* Use these to show how to load the here and palmerpenguins packages - libary() and require()

* Ginkgo data: use the 2021 ginkgo data to create a data.frame called ginkgo using:

### Require() function

This function used for loading the packages, which is already ready to use in R.

```{r}
require(here)
```

### Library() function

This function is about the all stored and downloaded packages in R
```{r}
library(palmerpenguins)
```

### Read.csv() function

this function allows csv files to open and rean in R. 

```{r}
ginkgo = data.frame(read.csv(here("data", "ginkgo_data_2021.csv")))
```



## 
## Data Structures

### c() function
The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type.
    * I can’t combine character and numeric types in the same call to c()

Here’s two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")
```

I can show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
## Typing the name of the vector into the console prints the contents
num_vec

## The print() function accomplishes the same task:
print(char_vec)
```

### print() function
This function print the contents of the vectors

```{r}
print(char_vec)
```

### length() function

This functions deals with the length of the vectors

```{r}
length(num_vec)
```
```{r}
length(char_vec)
```

### matrix() function
This function create a matrix from a given set of values to a no of rows and column. 

```{r}
matrix(num_vec, nrow = 3, ncol = 5, byrow = TRUE)
```

### data.frame() function
DataFrame is a data structure, organizes data into a table with rows and column. 
  * We can say ginkgo is a data frame and we can vizualize the ginkgo data as a table in R

```{r}
ginkgo
```
### Several useful functions to vizualize the specific data information from the DataFrame:

### nrow() function

```{r}
## Helps to vidualize the rows prsent in a data frame
nrow(ginkgo)
```

### ncol() function

```{r}
## Helps to vidualize the columns prsent in a data frame
ncol(ginkgo)
```

### dim() function

```{r}
## Helps to vidualize the rows and columns prsent in a data frame
dim(ginkgo)
```

## Subsetting

Useful function in R, which allow to filter only the specific part from the data set or data

### $

Selecting or mentioning (subsetting) a data set by the name

```{r}
head(ginkgo$petiole_length)
```

### [[]]

Selects data sets py position

```{r}
## Select the first row.
head(ginkgo[1,])
```
```{r}
## Select the third column.
head(ginkgo[,3])
```

### subset() function

subset use to select a certain subset from vectors, matrices or data frames

```{r}
## Recover all Adelie penguins data from the "palmerpenguins" dataset.
head(subset(penguins, species == "Adelie"))
```

## Numerical Data Exploration

### summary() function

This function provides the model summary

```{r}
summary(penguins)
```

### mean () function

This function provides the average (mean) of a data set

```{r}
# For observing the mean bill length
mean(penguins$bill_length_mm, na.rm = TRUE) 
```

### sd() function

This function produces the standard daviation from a dataset. For example we can calculate the standard deviation of bill length

```{r}
sd(penguins$bill_length_mm, na.rm = TRUE)
```

## Graphical Data Exploration

### Scatterplot()

  * Scatterplot can be generated usung the plott() function. Plot requires arguments as following: 
  
col: Deals with the colors. Useful functions for using the colors are- rgb() and colors() functions (color names and IDs)

pch: Assigns the types of points (circle, triangle, star etc.) within the plot

cex: Specifies the size of the points in the scatterplot

main: Title of the plot

xlab: Title of x-axis 

ylab: Title of y-axis

xlim: Controls the minimum and maximum values in the x-axis

ylim: Controls the minimum and maximum values in the y-axis

```{r}
plot(max_width ~ max_depth, data = ginkgo, main = "Maximum Depth and Width of Ginkgo Leaf", 
     xlab = "Maximum depth of leaf", ylab = "Maximum  width of leaf", 
     pch = 1, cex = 1,
     col = "blue4",
     xlim = c(20,150),
     ylim = c(20,150)
)
```

### hist() function
Creates histogram. 

```{r}
hist(penguins$flipper_length_mm, 
     breaks = 9, 
     main = "Penguin Flipper Length Histogram",
     xlab = "Flipper Length (mm)",
     col = "deepskyblue1")
```

### boxplot() function
1. a simple boxplot of ginkgo petiole lengths
```{r}
boxplot(
  ginkgo$petiole_length, 
  main = "Petiole Length of Ginkgo",
  ylab = "Petiole Length (mm)",
  col = "darkturquoise"
)
```

2. conditional boxplot of one of the continuous variables conditioned on the seeds_present column

```{r}
boxplot(
  data = ginkgo, max_width ~ seeds_present, 
  main = "Width of Ginkgo Leaf by Presense of Seeds",
  xlab = "Seeds Present", 
  ylab = "Maximum Width of Ginkgo (mm)", 
  names = c("Absent", "Present"),
  col = c("darksalmon", "darkseagreen2")
)
```

### 4-panel figure of histograms

```{r}
par(mfrow = c(2,2))

hist(ginkgo$max_width, 
     main = "Histogram of Ginkgo Leaf Width", 
     xlab = "Maximum Leaf Width (mm)", 
     col = "darkred")

hist(ginkgo$notch_depth, 
     main = "Histogram of Ginkgo Notch Depth", 
     xlab = "Notch Depth (mm)",
     col = "darkorchid1")

hist(ginkgo$petiole_length, 
     main = "Histogram of Ginkgo Petiole Length", 
     xlab = "Petiole Length(mm)",
     col = "brown1")

hist(ginkgo$max_depth, 
     main = "Histogram of Ginkgo Leaf Depth", 
     xlab = "Maximum Leaf Depth (mm)", 
     col = "darkgreen")
```

## Distribution Functions

  * Distribution functions are important to get idea about the probability of the specific values or the set of values. Distribution function is two types- 
    ** Normal Distribution
    ** Binomial Distribution

### Normal Distribution 

Known as the continuous probability distribution where values are placed in a symmetrical style and situated around the mean.

### dnorm() function

It's a probability density function of normal distribution

```{r}
dnorm(5, mean = 2, sd = 1)
## Probability of observing a value of 5, with a mean value 2 and a SD value 1.
```

### pnorm() function

Its a commulative distribution function of the normal distribution

```{r}
pnorm(q = 4, mean = 6, sd = 3)
## Probability of observing a value of 4 with a mean value 6 and a SD value 3.
```

### qnorm() function

Its a quantile function of normal distribution

```{r}
qnorm(p = 0.25, mean = 1, sd = 1)
## Probability one fourth (25%) with a mean and SD value 1.
```

### Binomial Distribution 

Known as the discrete probability distribution with two outcomes (number of successes or failure)

### dbinom() function
Its the probability density function for the binomial distribution

```{r}
dbinom(x = 5, size = 100, prob = 0.5)
## Probability of observing a value of 5 with a trial size of 100, and prob = 0.5
```

### pbinom() function

Cummulative distribution function for the binomial distribution

```{r}
pbinom(q = 4, size = 100, prob = 0.5)
## Probability of observing a value of less than 5 with a trial size of 100 and prob of 0.5
```

### qbinom() function

Quantile function for the binomial distribution

```{r}
qbinom(p = 0.5, size = 100, prob = 0.5)
## Probability of one fourth (25%) with a trial size of 100, and a prob. = 0.5
```



